---
title: 'Assignment # 1'
author: "Juan Gelvez-Ferreira"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r start, echo=FALSE}
rm(list=ls()) 

packages = c("tidyverse","readxl","ggplot2","haven")

package.check <- lapply(packages, FUN = function(x) {
if (!require(x, character.only = TRUE)) {
install.packages(x, dependencies = TRUE)
library(x, character.only = TRUE)
  
}})
```

## First assignment

*1.* Which command do you use to determine the type of an object?

Two alternatives. "type()" and/or typeof() to determines the type or storage mode of any object. 

```{r 1, echo=TRUE}
?type
?typeof
```

*2.* What is the type of vector A?
Using the command "typeof()", the vector A is a character. 
```{r 2, echo=TRUE}
A <- c("2", "3", "4", "5", "6", "7", "8")
typeof(A)
```

*3* Convert A into an integer vector
To convert A into an integer, I use the "as.integer()" function to create A1.

```{r 3, echo=TRUE}
?as.integer
A1 <- as.integer(A)
A1
typeof(A1)
```
*4*Create an integer vector B containing the numbers one through ten
```{r 4, echo=TRUE}
B <- (1:10)
B
typeof(B)
```
*5.* Create a new vector C from B which has the type "double"
```{r 5, echo=TRUE}
C<- as.double(B)
C
typeof(C)

```
*6*Change the third value of B to "3.5"
```{r 6, echo=TRUE}
B
B1<-replace(B,B==3,3.5)
B1
```

*7*Did this affect the type of B? How?
Yes, it moved from integer to double. Since "B" was containing the numbers one through ten, and the new B "B1" has a decimal, the type changed from a integer to a double. 
```{r 7, echo=TRUE}
typeof(B1)
```

#Reading in data

*8.* Read in the .dta version and store in an object called angell_stata
```{r 8, echo=TRUE}
angell_stata <- read_dta("data/angell.dta")
head(angell_stata)
```

*# 9.* Read in the .txt version and store it in an object called angell_txt
```{r 9, echo=TRUE}
angell_txt <- read.table("data\\angell.txt")
head(angell_txt)
```

*10.* Drop the first five observations in the angell_txt object
```{r 10, echo=TRUE}
angell_txt2<-tail(angell_txt, -5)
head(angell_txt2)
```

*11.* Select columns 2 and 3 of the agell_stata object and store them in a new object called angell_small
```{r 11, echo=TRUE}
angell_small<- angell_stata %>% 
  select(2,3)

head(angell_small)
```



*12.* Install the "MASS" package, load the package. Then, load the Boston dataset

```{r 12, warning=FALSE}
install.packages("MASS")
library(MASS)
head(Boston)
```

*13.* What is the type of the Boston object?
Using the "typeof" command, it's a list.
```{r 13, warning=FALSE}
typeof(Boston)
```


*14.* What is the class of the Boston object?
Boston is a dataframe. 
```{r 14, warning=FALSE}
class(Boston)
```


### Basic data summarizing and description
# 15. How many of the suburbs in the Boston data set bound the Charles river?

# 16. Do any of the suburbs of Boston appear to have particularly high crime rates? Tax rates?
# Pupil-teacher ratios? Comment on the range of each variable.

# 17. What is the median pupil-teacher ratio among the towns in this data set that
# have a per capita crime rate larger than 1 ?


### Functions
# 18. Write a function that calculates the squareroot of an integer


# 19. Write a function that calculates 95% confidence intervals for a point estimate.
# The function should be called "my_CI"
# When called with "my_CI(2, 0.2)", the output of the function should read
# "The 95% CI upper bound of point estimate 2 with standard error 0.2 is 2.392. The lower bound is 1.608."
# Note: the function should take a point estimate and its standard error as arguments
# You may use the formula for 95% CI: point estimate +/- 1.96*standard error)
# Pasting text in R: paste() and paste0()

# 20. Write a function that converts all negative numbers in the following dataset into NA
# Use as little code as possible and try to avoid code repetition
set.seed(1002)
df <- data.frame(replicate(10, sample(c(1:10, c(-99,-98,-5)), 6, rep = TRUE)))
names(df) <- letters[1:6]
df



# 21. Use your function to convert all negative numbers in the dataset into NA without changing the class of the object



# 22. Change the function you wrote above such that it turns any negative number into NA!


