---
title: "Fundamentals of Computing and Data Display"
subtitle: "Interactive Graphs"
author: Christoph Kern <br> <font size="2">Thanks to Georg-Christoph Haas (IAB)!</font>
output: html_notebook
---

## Setup

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(ggvis)
library(shiny)
library(gapminder)
```

## Data

In this notebook we use the gapminder dataset. The general idea is that we want to build a lightweight version of the interactive GDP versus life expectancy graph with ggvis().

https://www.gapminder.org/tools

Before we start with creating graphs, we take a look at the data.

```{r}
head(gapminder)
summary(gapminder)

?gapminder
```

## Task 1: Basic graph 

We'll create our graph step by step, starting with a static scatterplot that maps GDP per capita on the x-axis and life expectancy on the y-axis. We map the variable `continent` to fill and set size equivalent to `pop`.

```{r}
gapminder %>%
  filter(year == 2007) %>%
  ggvis(x = ~gdpPercap,
        y = ~lifeExp,
        fill = ~continent,
        size := ~pop/250000,
        opacity := 0.5) %>%
  layer_points(stroke := "white")
```

## Task 2: Adding a hover option

Now we take our static graph from task 1 and add a hover option to it (`add_tooltip()`).

```{r}
gapminder %>%
  filter(year == 2007) %>%
  ggvis(x = ~gdpPercap,
        y = ~lifeExp,
        fill = ~continent,
        size := ~pop/250000,
        opacity := 0.5) %>%
  layer_points(stroke := "white") %>%
  add_tooltip(function(data){paste0(
    "lifeExp: ", round(data$lifeExp), "<br>",
    "gdpPercap: ", round(data$gdpPercap)) })
```

## Task 3: Extending the hover option

Now we add country to the function `addtooltip()`. This requires to add an additional `layer_points()` that allows to access the `country` variable.

```{r}
gapminder %>%
  filter(year == 2007) %>%
  ggvis(x = ~gdpPercap,
        y = ~lifeExp,
        fill = ~continent,
        size := ~pop/250000,
        opacity := 0.5) %>%
  layer_points(stroke := "white") %>%
  layer_points(fill := ~country, 
               opacity = 0) %>%
  add_tooltip(function(data){paste0(
    "country: ", data$country, "<br>",
    "lifeExp: ", round(data$lifeExp), "<br>",
    "gdpPercap: ", round(data$gdpPercap)) })
```

## Task 4: Choosing a continent

Our basic graph only takes data from 2007. Still, the graph seems crowded and it would be nice to choose which continents are shown. We can solve this with `dplyr`s `filter()` function and `input_checkboxgroup()`. 

```{r}
gapminder %>%
  filter(year == 2007) %>%
  ggvis(x = ~gdpPercap,
        y = ~lifeExp,
        fill = ~continent,
        size := ~pop/250000,
        opacity := 0.5) %>%
  filter(continent %in% 
           eval(input_checkboxgroup(
             c("Americas", "Africa", "Asia", "Europe", "Oceania"), 
             selected = c("Americas", "Africa", "Asia", "Europe", "Oceania"))))
```

## Task 5: Adding a slider for years

In our previous graphs we only use data from 2007. Here we add a slider with the `input_slider()` function to select between years.

```{r}
gapminder %>%
  ggvis(x = ~gdpPercap,
        y = ~lifeExp,
        fill = ~continent,
        size := ~pop/250000,
        opacity := 0.5) %>%
  filter(year == eval(input_slider(min = 1952, max = 2007, value = 1952, step = 5, animate = T, sep = ""))) %>%
   layer_points(stroke := "white")
```

## Task 6: Adding fixed axes

The next task is to fix the scales so that we can observe changes over years. First, we define an invisible minimum and maximum point with `layer_points()`. Next, we use the function `add_axis()` to choose reasonable levels for the axis. 

```{r}
gapminder %>%
  ggvis(x= ~gdpPercap,
        y= ~lifeExp,
        fill= ~continent,
        size:=~pop/250000,
        opacity:=0.5) %>%
  filter(year == eval(input_slider(min = 1952, max = 2007, value = 1952, step = 5, animate = T, sep = ""))) %>%
  layer_points(stroke := "white") %>%
  layer_points(x = min(gapminder$gdpPercap), y = min(gapminder$lifeExp), opacity = 0) %>%
  layer_points(x = max(gapminder$gdpPercap), y = max(90), opacity = 0) %>%
  add_axis("y", values = seq(30 , 90, by = 10)) %>%
  add_axis("x", values = c(25000, 50000, 75000, 100000))
```

## Task 7: Slider axes (zoom)

In the next task we take our fixed axes and make them interactive again by adding sliders.

```{r}
y_axis <- input_slider(min = 30, max = 90, value = c(30, 90))
x_axis <- input_slider(min = 0, max = 100000, value = c(1000, 100000), step = 1000)

gapminder %>%
  filter(year == 2007) %>%
  ggvis(x = ~gdpPercap,
        y = ~lifeExp,
        fill = ~continent,
        size := ~pop/250000,
        opacity := 0.5) %>%
  layer_points(stroke := "white") %>%
  scale_numeric("y", domain = y_axis, clamp = T) %>%
  scale_numeric("x", domain = x_axis, clamp = T)
```

## Task 8: Your own ideas

In this chunk you can add your own ideas. Try to add another interactive component to the graph.

```{r}
gapminder %>%
  filter(year == 2007) %>%
  ggvis(x = ~gdpPercap,
        y = ~lifeExp,
        fill = ~continent,
        size := ~pop/250000,
        opacity := 0.5) %>%
  layer_points(stroke := "white") 
```

## Task 9: Putting it all together

Finally, we can combine all pieces from above and create our final interactive graph. (The viewer may be too small to display this graph. Open it in the browser.)

```{r}
y_axis <- input_slider(min = 30, max = 90, value = c(30, 90), label= "Range of Life Expectancy")
x_axis <- input_slider(min = 0, max = 100000, value = c(1000, 100000), step = 1000, label = "Range GDP per Capita")

gapminder %>%
  ggvis(x = ~gdpPercap,
        y = ~lifeExp,
        fill = ~continent,
        size := ~pop/250000,
        opacity := 0.5) %>%
  filter(continent %in% 
           eval(input_checkboxgroup(
             c("Americas", "Africa", "Asia", "Europe", "Oceania"), 
             selected = c("Americas", "Africa", "Asia", "Europe", "Oceania"),
             label = "Choose a Continent"))) %>%
  filter(year %in% eval(input_slider(min = 1952, max = 2007, value = 1952, step = 5, animate = T, label = "Year (Press Play)", sep = ""))) %>%
  layer_points(stroke := "white") %>%
  layer_points(fill := ~country, 
               opacity = 0) %>%
  add_tooltip(function(data){paste0(
    "country: ", data$country, "<br>",
    "lifeExp: ", round(data$lifeExp), "<br>",
    "gdpPercap: ", round(data$gdpPercap)) }) %>%
  scale_numeric("y", domain = y_axis, clamp = T) %>%
  scale_numeric("x", domain = x_axis, clamp = T) %>%
  add_axis("x", title_offset = 40, title = "GDP per Capita in US-Dollar (inflation adjusted)", properties = axis_props(title = list(fontSize = 15))) %>%
  add_axis("y", title_offset = 40, title = "Life Expectancy in years", properties = axis_props(title = list(fontSize = 15)))
```
